---
title: "ADFG_report_2020"
author: "Lia Domke"
date: "2/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For our ADFG permit it requires report the number of each species that were collected during that reporting year. 

Species groups we collected in 2020:
1. fish species - seine
2. invertebrate species - seine
3. crabs - pots
4. fish - pots
5. eelgrass - habitat 
6. eelgrass - eelgrass wasting disease
7. kelp - habitat

Based on WR script 

# Alaska Department of Fish and Game Fisheries Resource permit reporting
This script will take raw "tall" data, from multiple parent sources if necessary, and format it for the FRP reporting form.
```{r libraries}
library(dplyr)
library(tidyr)
library(readxl)
library(lubridate)
```


## Data
```{r data}
# Eelgrass Site
site <- read.csv("../APECS Master repository/APECS Master repo/ALL_DATA/Site_metadata_allyears.csv", header = TRUE,
                 stringsAsFactors = FALSE)
# Eelgrass Biometrics (lab data)
eg.bio <- read.csv("../APECS Master repository/APECS Master repo/ALL_DATA/Lia_fish/Eelgrass_biomatrics_2019.csv",
                   stringsAsFactors = FALSE, header = TRUE)
# Habitat data
hab_quad <- read_xlsx("../APECS Master repository/APECS Master repo/ALL_DATA/Lia_fish/Habitat_2019_RAW_2-21-2020.xlsx", 
                      sheet = 1)
hab_plant_lng <- read_xlsx("../APECS Master repository/APECS Master repo/ALL_DATA/Lia_fish/Habitat_2019_RAW_2-21-2020.xlsx",
                           sheet = 2)
hab_plant_biom <- read_xlsx("../APECS Master repository/APECS Master repo/ALL_DATA/Lia_fish/Habitat_2019_RAW_2-21-2020.xlsx",
                            sheet = 3)
# Eelgrass wasting disease
# Crab Pot
crab.dat <- read.csv("../APECS Master repository/APECS Master repo/ALL_DATA/Lia_fish/crab_data_2019.csv", 
                     header = TRUE, stringsAsFactors = FALSE)
# Beach Seine
fish.dat <- read.csv("../APECS Master repository/APECS Master repo/ALL_DATA/Lia_fish/FishSeine_2019_RAW_10-4-19.csv",
                     stringsAsFactors = FALSE, header = TRUE)

fish.dat$unmeasured <- as.numeric(fish.dat$unmeasured)
```

## Site info
Subset by the sites surveyed in 2019
```{r}
site.19 <- site %>%
  subset(site_2019 %in% fish.dat$site) %>%
  unite(universal_code, bay_code:bay_sample, sep = "_", remove = FALSE)

site <- site %>%
  unite(universal_code, bay_code:bay_sample, sep = "_", remove = FALSE)
```

## Kelp biomass

```{r}
# adjust sites names so that we can convert to universal code
hab_quad$site <- as.factor(hab_quad$site)
levels(hab_quad$site)[levels(hab_quad$site)=="North Fish Egg - kelp"] <- "North Fish Egg-Kelp"
levels(hab_quad$site)[levels(hab_quad$site)=="Baker Island - kelp"] <- "Baker Island - kelp "
unique(hab_quad$site)
# join by quadratID 
kelp_biom <- left_join(hab_plant_biom, hab_quad, by = "quadrat_ID") %>%
  dplyr::select(species_by_quadrat_ID:habitat, date) %>%
  left_join(site.19, by = c("site" = "site_2019")) %>%
  select(c(site, universal_code, species, weight, date))
  
kelp_sum <- kelp_biom %>%
  group_by(universal_code, date, site, species) %>%
  dplyr::summarise(weight_g = sum(weight)) %>%
  mutate(sp = species) %>%
  ungroup() %>%
  dplyr::select(c(sp, universal_code, weight_g, site, date))

kelp_sum$collection_method <- "hand"

kelp <- kelp_sum
kelp <- mutate(kelp, date = ymd(date))
```


## Eelgrass
Our protocol for eelgrass collections for HABITAT data included 15 shoots at 5 quadrats at all eelgrass sites (20 sites)
for a total of 1500 shoots
```{r n eelgrass}
# get right site info
site.eg <- site.19 %>%
  filter(habitat == "eelgrass") %>%
  subset(site_2019 %in% fish.dat$site) 

eg <- data.frame("site" = unique(site.eg$site_2019), "sp" = rep("Zostera marina", 20), "collection_method" = "hand", "count" = 75, "universal_code" = unique(site.eg$universal_code))

eg <- left_join(eg, hab_quad, by = "site") %>%
  dplyr::select(site, sp, collection_method, count, universal_code, date) %>%
  distinct()

# fix dates
eg[3, "date"] <- "2019-05-08"
eg[5, "date"] <- "2019-05-22"
eg[6, "date"] <- "2019-06-17"
eg[7, "date"] <- "2019-05-18"
eg[9, "date"] <- "2019-06-06"
eg[11, "date"] <- "2019-06-16"
eg[12, "date"] <- "2019-05-21"
eg[13, "date"] <- "2019-05-23"
eg[18, "date"] <- "2019-05-20"
eg[20, "date"] <- "2019-05-10"

eg$date <- as.Date(eg$date)

```
 
## Eelgrass wasting disease
At 5 sites 132 shoots were taken, at the 6th site only 120 were taken
```{r}
waste.site <- c("NATZ_A", "GUKT_A", "NFEI_A", "REFU_A", "NOSK_A", "SHIN_A")
waste.date <- c("2019-07-17", "2019-07-15", "2019-07-14", "2019-07-19", "2019-07-16","2019-07-18")
eg.waste <- data.frame("site" = waste.site, "date" = waste.date, sp = rep("Zostera marina", 6), "collection_method" = "hand", "count" = 132)
# fix site where we collected fewer eelgrass blades
eg.waste[4,5] <- "120"

eg.waste <- mutate(eg.waste, date = ymd(date))
eg.waste$count <- as.numeric(eg.waste$count)
```


## Grazers - small inverts
This will come from the biometrics data
```{r n grazers}
## Summarise ##
eg.bio <- eg.bio %>%
  mutate(date = mdy(collection_date))

graz <- data.frame(
  eg.bio %>% 
    group_by(place_name, collection_date) %>% 
    summarise(Idothea_rascata = sum(idotea_count, na.rm = TRUE),
              Gammarid_amphipod = sum(amphipod_count, na.rm = TRUE),
              Caprella_alaskana = sum(caprellid_count, na.rm = TRUE),
              Lottia_sp = sum(limpet_count, na.rm = TRUE),
              Littorina_sp = sum(gastropod_count, na.rm = TRUE)) %>%
    mutate(site = place_name) %>%
    mutate(date = mdy(collection_date)) %>%
    ungroup() %>%
    dplyr::select(-c(place_name, collection_date))
  )
## Convert to tall ##
graz <- graz %>%
  gather(sp, count, Idothea_rascata:Littorina_sp) %>%
  left_join(site.eg, by = c("site" = "place_name")) %>%
  dplyr::select(c(site, sp, count, universal_code, date))
## Add colelction method ##
graz$collection_method <- "hand"
colnames(graz)[3] <- "count"
```

## Crabs
This data will come from the crab pot data. We also caught fish in pots so that will be extracted here and added to fish data later.

First you should check the place_names and make sure that it is consistent with what is on the site_metadata. Convert to universal name and then convert back
```{r}
crab.dat2 <- crab.dat %>%
  unite(universal_code, bay_code:bay_sample) %>%
  left_join(site, by = "universal_code") %>%
  mutate(place_name = place_name.x) %>%
  mutate(date = ymd(YYYYMMDD_pulled)) %>%
  dplyr::select(c(universal_code, Sp_ScientificName, taxa, place_name, date))
```


```{r n crabs}
## Crabs from pots ##
crab <- data.frame(
  crab.dat2 %>% 
    filter(taxa == "Decapoda") %>% 
    group_by(place_name, Sp_ScientificName, universal_code, date) %>% 
    summarise(count = n()) %>%
    mutate(site = place_name) %>%
    mutate(sp = Sp_ScientificName) %>%
    ungroup() %>%
    dplyr::select(-c(place_name, Sp_ScientificName))
)

## Fish from pots ##
fish.crab <- data.frame(
  crab.dat2 %>% 
    filter(taxa == "Vertebrata") %>% 
    group_by(place_name, Sp_ScientificName, universal_code, date) %>% 
    summarise(count = n()) %>%
    mutate(site = place_name) %>%
    mutate(sp = Sp_ScientificName) %>%
    ungroup() %>%
    dplyr::select(-c(place_name, Sp_ScientificName))
)
```

## Fish
This data will come from the beach seine data. We also caught crabs in the beach seine so that will be extracted here and added to the crab derived above.
```{r n fish}
fish.dat$abundance <- as.numeric(ifelse(is.na(fish.dat$fork_length), paste(fish.dat$unmeasured), 1))

## Fish from seine ##
fish.sum <- fish.dat %>%
  filter(taxon == "fish") %>%
  left_join(site, by = c("site" = "site_2019")) %>%
  unite(universal_code, bay_code:bay_sample) %>%
  mutate(date = mdy(date)) %>%
  group_by(site, species_scientific, universal_code, date) %>%
  dplyr::summarise(count = sum(abundance)) %>%
  ungroup() %>%
  mutate(sp = species_scientific) %>%
  dplyr::select(-species_scientific)


## Crabs from seine ##
crab.fish <- data.frame(
  fish.dat %>%
    filter(taxon == "invertebrate") %>% 
    left_join(site, by = c("site" = "site_2019")) %>%
    unite(universal_code, bay_code:bay_sample) %>%
    mutate(date = mdy(date)) %>%
    group_by(site, species_scientific, universal_code, date) %>% 
    summarise(count = n()) %>%
    ungroup() %>%
    mutate(sp = species_scientific) %>%
    dplyr::select(-species_scientific)
)
```

## Bring it all together now
Get all the summaries above into one nice happy table! Including site identifier information.
```{r}
## All the crabs ##
crab$collection_method <- "crab pot"
crab.fish$collection_method <- "beach seine"
crab <- bind_rows(crab.fish, crab)
## All the fish ##
fish.crab$collection_method <- "crab pot"
fish.sum$collection_method <- "beach seine"
fish <- bind_rows(fish.sum, fish.crab)
fish[is.na(fish)] <- 0
## Final Combine ##


counts <- bind_rows(crab, fish, eg, graz, eg.waste, kelp)

sites_19 <- unique(counts$universal_code) 

dat <- site %>%
  dplyr::select(c(place_name, universal_code, latitude, longitude)) %>%
  subset(universal_code %in% sites_19)

dat2 <- left_join(dat, counts)
dat2$Life_stage <- "adult/non-larval juveniles"



## Just group by species ##
sp <- dat2 %>% 
 group_by(sp) %>% 
summarise(total = sum(count), total_weight = sum(weight_g))
## Just fish for IACUC ##
#fish.sp <- fish %>% 
 # group_by(sp) %>% 
  #summarise(total = sum(count))
#fish.sp$total_retained <- 0
```

## Export
```{r}
write.csv(dat2, "Data/ADFG_collections_CF_19_067.csv", row.names = FALSE)
#write.csv(fish.sp, "../../IACUC/IACUC_collections_2017.csv", row.names = FALSE)
```