---
title: "ADFG_reporting_2020"
author: "Lia Domke"
date: "1/21/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For our ADFG permit it requires report the number of each species that were collected during that reporting year. 

Species groups we collected in 2020:
1. fish species - seine
2. invertebrate species - seine
3. crabs - pots (multiple depths this time)
4. fish - pots (multiple depths this time)
5. eelgrass - biometics only no hab
6. eelgrass - eelgrass wasting disease


Based on WR script 

# Alaska Department of Fish and Game Fisheries Resource permit reporting
This script will take raw "tall" data, from multiple parent sources if necessary, and format it for the FRP reporting form.
```{r libraries}
library(dplyr)
library(tidyr)
library(readxl)
library(lubridate)
```


## Data
```{r data}
# Eelgrass Site
site <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A87744ced-530b-4bae-809e-ff1012c7ae72"), header = TRUE, stringsAsFactors = FALSE)

# Eelgrass Biometrics (lab data)
eg.bio <- read.csv("Data/eelgrass_biometrics_2020_RAW_unchecked.csv",
                   stringsAsFactors = FALSE, header = TRUE)

# Eelgrass wasting disease (count only, no biomass)

# Crab Pot (multiple depths)
crab.dat.near <- read.csv("Data/nearshore_crab_data_2020_RAW_unchecked.csv", 
                     header = TRUE, stringsAsFactors = FALSE)
crab.dat.30depth <- read.csv("Data/30ft_depth_crab_data_2020_RAW_unchecked.csv", 
                             header = TRUE, stringsAsFactors = FALSE)
crab.dat.depth <- read.csv("Data/Depth_gradient_crab_RAW_2020_unchecked.csv", 
                           header = TRUE, stringsAsFactors=FALSE)

# Beach Seine
fish.dat <- read.csv("Data/fish_beach_seine_2020_RAW_unchecked.csv",
                     stringsAsFactors = FALSE, header = TRUE)

fish.dat$unmeasured <- as.numeric(fish.dat$unmeasured)

# add in last years approved IACUC species list to  make sure we were approved to collect said fish
iacuc.sp.list <- read_xlsx("../../Permits:Licensing/IACUC/Field_season_2020/Updated_IACUC_species_2020.xlsx",
                           sheet = 1)
```

## Site info
Subset by the sites surveyed in 2020
```{r}
site <- site %>%
  unite(universal_code, bay_code:bay_sample, sep = "_", remove = FALSE)
```


## Eelgrass wasting disease
20 shoots were taken per transect. 6 transects per site.
```{r}
waste.uni.code <- c("NATZ_A", "GUKT_A", "NFEI_A", "REFU_A", "NOSK_A", "SHIN_A")
waste.date <- c("2020-07-21", "2020-07-24", "2020-08-01", "2020-07-23", "2020-08-02","2020-07-19")
eg.waste.df <- data.frame("site" = waste.uni.code, "date" = waste.date, sp = rep("Zostera marina", 6), "collection_method" = "hand", "count" = 120)

eg.waste.df <- mutate(eg.waste.df, date = ymd(date))
eg.waste.df$count <- as.numeric(eg.waste.df$count)

eg.waste <- eg.waste.df %>%
  left_join(site, by = c("site" = "universal_code")) %>%
  rename("universal_code" = "site") %>%
  dplyr::select(c(place_name, universal_code, date, sp, collection_method, count))
  
```


## Grazers - small inverts
This will come from the biometrics data
```{r n grazers}
## Summarise ##
eg.bio <- eg.bio %>%
  mutate(date = mdy(collection_date))

graz.df <- data.frame(
  eg.bio %>% 
    group_by(place_name, collection_date) %>% 
    summarise(Idothea_rascata = sum(idotea_count, na.rm = TRUE),
              Gammarid_amphipod = sum(amphipod_count, na.rm = TRUE),
              Caprella_alaskana = sum(caprellid_count, na.rm = TRUE),
              Lottia_sp = sum(limpet_count, na.rm = TRUE),
              Littorina_sp = sum(gastropod_count, na.rm = TRUE)) %>%
    mutate(site = place_name) %>%
    mutate(date = mdy(collection_date)) %>%
    ungroup() %>%
    dplyr::select(-c(place_name, collection_date))
  )
## Convert to tall ##
graz <- graz.df %>%
  gather(sp, count, Idothea_rascata:Littorina_sp) %>%
  left_join(site, by = c("site" = "place_name")) %>%
  rename(place_name = site) %>%
  dplyr::select(c(place_name,universal_code, sp, count, date))
## Add colelction method ##
graz$collection_method <- "hand"
colnames(graz)[4] <- "count"
```

## Crabs
This data will come from the crab pot data. We also caught fish in pots so that will be extracted here and added to fish data later.

First you should check the place_names and make sure that it is consistent with what is on the site_metadata. Convert to universal name and then convert back

```{r}

# Combine two of the data files
crab.combined <- crab.dat.depth %>%
  full_join(crab.dat.30depth) %>%
  full_join(crab.dat.near)


# fix names
crab.combined$bay_code <- as.factor(crab.combined$bay_code)
levels(crab.combined$bay_code)[levels(crab.combined$bay_code)=="GOATS"] <- "GOAT"
levels(crab.combined$bay_code)[levels(crab.combined$bay_code)=="KLWK"] <- "KLWA"
levels(crab.combined$bay_code)[levels(crab.combined$bay_code)=="NPASS"] <- "NPAS"
levels(crab.combined$bay_code)[levels(crab.combined$bay_code)=="GUTK"] <- "GUKT"

anti_join(crab.combined,site, by = "bay_code") %>%
  select(bay_code) %>%
  distinct() # none missing 
```


```{r}
# create a diff site that just has one bay code for site id for sites that happened within the bay but not at a specific bed level. 
bay_sites <- site %>%
  select(bay_code, place_name) %>%
  distinct() 

crab.dat2 <- crab.combined %>%
  #unite(universal_code, bay_code:bay_sample) %>%
  left_join(bay_sites, by = "bay_code") %>%
  mutate(date = ymd(YYYYMMDD_pulled)) %>%
  dplyr::select(c(bay_code, Sp_ScientificName, Taxa, place_name, date, carapace_width_mm, sex, fork_length_mm, unmeasured))

```


```{r n crabs}
## Crabs from pots ##
crab <- data.frame(
  crab.dat2 %>% 
    filter(Taxa == "invertebrate") %>% 
    group_by(place_name, Sp_ScientificName, bay_code, date) %>% 
    summarise(count = n()) %>%
    mutate(sp = Sp_ScientificName) %>%
    ungroup() %>%
    dplyr::select(-c(bay_code, Sp_ScientificName))
)

## Fish from pots ##
fish.crab <- data.frame(
  crab.dat2 %>% 
    filter(Taxa == "fish") %>% 
    group_by(place_name, Sp_ScientificName, bay_code, date) %>% 
    summarise(count = n()) %>%
    mutate(sp = Sp_ScientificName) %>%
    ungroup() %>%
    dplyr::select(-c(bay_code, Sp_ScientificName))
)
```

## Fish
This data will come from the beach seine data. We also caught crabs in the beach seine so that will be extracted here and added to the crab derived above.
```{r n fish}
# lets drop the last couple colmns cause they're associated w/ data checking
fish.dat <- fish.dat[,-c(22:28)]
fish.dat$place_name <- as.factor(fish.dat$place_name)
levels(fish.dat$place_name)[levels(fish.dat$place_name)=="Natzuhini"] <- "Natzuhini Bay"
fish.dat$place_name <- as.character(fish.dat$place_name)

fish.dat <- fish.dat %>%
  unite(unmeasured, unmeasured:unmeasured_sm, na.rm = TRUE, remove = FALSE)

fish.dat$abundance <- as.numeric(ifelse(is.na(fish.dat$fork_length), paste(fish.dat$unmeasured), 1))

## Fish from seine ##
fish.sum <- fish.dat %>%
  filter(taxon == "vertebrate") %>%
  left_join(site, by = c("place_name", "habitat", "bay_code", "bay_sample")) %>%
  unite(universal_code, bay_code:bay_sample) %>%
  mutate(date = mdy(date)) %>%
  group_by(place_name, species_scientific, universal_code, date) %>%
  dplyr::summarise(count = sum(abundance)) %>%
  ungroup() %>%
  mutate(sp = species_scientific) %>%
  dplyr::select(-c(species_scientific))


## Crabs from seine ##
crab.fish <- data.frame(
  fish.dat %>%
    filter(taxon == "invertebrate") %>% 
    left_join(site, by = c("habitat", "bay_code", "bay_sample", "place_name")) %>%
    unite(universal_code, bay_code:bay_sample) %>%
    mutate(date = mdy(date)) %>%
    group_by(place_name, species_scientific, universal_code, date) %>% 
    summarise(count = n()) %>%
    ungroup() %>%
    mutate(sp = species_scientific) %>%
    dplyr::select(-c(species_scientific))
)
```

## Bring it all together now
Get all the summaries above into one nice happy table! Including site identifier information.
```{r}
## All the crabs ##
crab$collection_method <- "crab pot"
crab.fish$collection_method <- "beach seine"
crab <- bind_rows(crab.fish, crab)
## All the fish ##
fish.crab$collection_method <- "crab pot"
fish.sum$collection_method <- "beach seine"
fish <- bind_rows(fish.sum, fish.crab)
fish[is.na(fish)] <- 0
## Final Combine ##


counts <- bind_rows(crab, fish, graz, eg.waste) %>%
  dplyr::select(-universal_code)

# get lat/long for crab pot sites
crab.dat.near$bay_code <- as.factor(crab.dat.near$bay_code)
levels(crab.dat.near$bay_code)[levels(crab.dat.near$bay_code)=="GOATS"] <- "GOAT"
levels(crab.dat.near$bay_code)[levels(crab.dat.near$bay_code)=="KLWK"] <- "KLWA"
levels(crab.dat.near$bay_code)[levels(crab.dat.near$bay_code)=="NPASS"] <- "NPAS"
levels(crab.dat.near$bay_code)[levels(crab.dat.near$bay_code)=="GUTK"] <- "GUKT"
crab.dat.near$bay_code <- as.character(crab.dat.near$bay_code)
sites_20 <- crab.dat.near %>%
  unite(universal_code, bay_code:bay_sample) %>%
  dplyr::select(universal_code) %>%
  distinct() %>%
  left_join(site, by = "universal_code") %>%
  dplyr::select(c(universal_code, place_name, latitude, longitude))

dat <- full_join(sites_20, counts)
dat$Life_stage <- "adult/non-larval juveniles"



## Just group by species ##
sp <- dat %>% 
 group_by(sp) %>% 
summarise(total = sum(count))

## Just fish for IACUC ##
# add back in common name
sp_list <- fish.dat %>%
  filter(taxon == "vertebrate") %>%
  dplyr::select(species_common, species_scientific) %>%
  distinct()

# get number of fish caught for each species
fish.sp <- fish %>% 
  filter(collection_method == "beach seine") %>% # only  beach seine data
  group_by(sp) %>%
  summarise(total = sum(count)) %>%
  left_join(sp_list, by = c("sp" = "species_scientific")) %>%
  relocate(species_common, sp, total)
  
fish.sp$total_retained <- 0
```

## Export
```{r}
write.csv(dat, "Data/Species_by_site_2020.csv", row.names = FALSE)
write.csv(sp, "Data/Species_2020.csv", row.names = FALSE)
#write.csv(fish.sp, "../../Permits:Licensing/IACUC/IACUC_collected_in_2020.csv", row.names = FALSE)
```