---
title: "IACUC_collected_2022"
author: "Lia Domke"
date: "3/22/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(readxl)
```

We need to report to IACUC the species and how many individuals of each species we handled/caught in 2022

Read in data
```{r}
seine22 <- read.csv("../APECS_DM/Data/Fish_Seine/fish_mass_2022_derived.csv") #make sure its the most up to date.
# this one has already been cleaned by APECS_EM so thats great

# read in species list information from knb
species <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A08b8bcd1-3451-4a3a-a9ee-f3ad8b1a437e"), header = TRUE, stringsAsFactors = FALSE)


# add in last years approved IACUC species list to  make sure we were approved to collect said fish
iacuc.sp.list <- read_xlsx("../../Permits:Licensing/IACUC/Field_season_2022/iacuc_permited_species_2022.xlsx",
                           sheet = 1)
```

Take a look at the data format
```{r}
glimpse(seine22)

unique(seine22[c("sp_code", "species_common", "species_scientific")])
unique(seine22[c("bay_id","place_name", "date", "start_time", "end_time", "slope", "tide_height", "tide_time", "habitat")])
```

Basic cleaning
```{r}
fish <- seine22 %>%
  filter(taxon == "fish") # remove invertebrate catch

```

Calculate number of individuals caught of each species over the season
```{r}
fish.sp <- fish %>% 
  group_by(species_scientific) %>%
  summarise(total = sum(abundance)) %>%
  left_join(species, by = c("species_scientific" = "Sp_ScientificName"), ("species_common" = "Sp_CommonName")) %>%
  relocate(Sp_CommonName, species_scientific, total)

fish.sp$total_retained <- 0
```

Lets see the max catch from the IACUC forms
```{r}
glimpse(iacuc.sp.list)
```

Looks pretty good, i.e. that we don't have any species we didn't account for or touched too many fish
```{r}
#write.csv(fish.sp, "../../Permits:Licensing/IACUC/Field_season_2022/IACUC_collected_2022.csv")
```


For the ADFG report: 
how much fish in mass did we catch
```{r}
glimpse(fish)

mass.by.sp <- fish %>%
  group_by(bay_id, sp_code, species_scientific) %>%
  dplyr::summarise(total_mass = sum(mass_g))

range(mass.by.sp$total_mass)

mass.by.site <- fish %>%
  group_by(bay_id) %>%
  dplyr::summarise(total_mass = sum(mass_g)/100,
                   avg_mass = mean(mass_g)/100)

range(mass.by.site$total_mass)
range(mass.by.site$avg_mass)

fish %>%
  group_by(bay_id) %>%
  summarise(avg_mass = mean(mass_g)/100)

unique(fish[c("sp_code", "species_common", "species_scientific")])

sp.counts <- fish %>%
  group_by(species_common, species_scientific) %>%
  dplyr::summarise(counts = sum(abundance)) %>%
  arrange(-counts)

fish %>%
  summarise(counts = sum(abundance))

# used these above metrics to calculate the things required for the ADFG research report
```


