---
title: "IACUC_collected_2021"
author: "Lia Domke"
date: "9/3/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(readxl)
```

We need to report to IACUC the species and how many individuals of each species we handled/caught in 2021

Read in data
```{r}
seine21 <- read.csv("../2021_data/fish_beach_seine_2021_MASTER_9-3-21.csv", stringsAsFactors = FALSE) #make sure its the most up to date.
# 9/3/21 has been entered and initially data checked. There is some discrepancy that hasn't been resolved yet for Shinaku small shiners (tbd exactly how many we caught). So there will be an offset of between 5 - 250 shiners

# read in species list information from knb
species <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A08b8bcd1-3451-4a3a-a9ee-f3ad8b1a437e"), header = TRUE, stringsAsFactors = FALSE)


# add in last years approved IACUC species list to  make sure we were approved to collect said fish
iacuc.sp.list <- read_xlsx("../../Permits:Licensing/IACUC/Field_season_2021/IACUC_species_2021.xlsx",
                           sheet = 1)
```

Take a look at the data format
```{r}
glimpse(seine21)
```

Basic cleaning
```{r}
fish <- seine21 %>%
  dplyr::select(-c(FL_checking, same.)) %>% # remove data checking columns
  unite("site_ID", bay_code:bay_sample) %>% # create a single column for site code
  filter(taxon == "vertebrate") # remove invertebrate catch

fish$abundance <- as.numeric(ifelse(is.na(fish$unmeasured), 
       ifelse(is.na(fish$unmeasured_sm), 1, paste(fish$unmeasured_sm)), 
       paste(fish$unmeasured))) # create abundance column that prints either a 1 or actual number of fish     caught
```

Calculate number of individuals caught of each species over the season
```{r}
fish.sp <- fish %>% 
  group_by(species_scientific) %>%
  summarise(total = sum(abundance)) %>%
  left_join(species, by = c("species_scientific" = "Sp_ScientificName"), ("species_common" = "Sp_CommonName")) %>%
  relocate(Sp_CommonName, species_scientific, total)

fish.sp$total_retained <- 0
```

Lets see the max catch from the IACUC forms
```{r}
glimpse(iacuc.sp.list)
```

Looks pretty good, i.e. that we don't have any species we didn't account for or touched too many fish
```{r}
#write.csv(fish.sp, "../../Permits:Licensing/IACUC/Field_season_2021/IACUC_collected_2021.csv")
```

Looking at average eelgrass biomass, epiphyte load, and grazer load from 2021
```{r}
eel.biom.dens21 <- read.csv("../Chp3_SeascapeDynamics/Data/eelgrass_biom_2021.csv", stringsAsFactors = FALSE, header = TRUE)


eel.biom.dens21 %>%
  group_by(site) %>%
  dplyr::summarise(avg_biom_g_m2 = mean(total_biomass_g))
```

