---
title: "Kelp_sp_2020"
author: "Lia Domke"
date: "2/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In 2019 we collected a lot of kelps that we didn't anticipate. In order to update our ADFG permit for 2020, lets create a csv with species names and approximate abundances based on how many kelp sites we'll have.

1. read in data
2. subset to just kelp data
3. approximate biomass for 2020 sites based on 2019
  Note that in 2019 we sampled 5 kelp sites and in 2020 we will sample 6 kelp sites

# Libraries
```{r message=TRUE}
require(tidyverse)
require(readxl)
require(lubridate)
```

  
# 1. Data
```{r message=TRUE}
hab_quad <- read_xlsx("../APECS Master repository/APECS Master repo/ALL_DATA/Lia_fish/Habitat_2019_RAW_2-21-2020.xlsx", 
                      sheet = 1)
hab_plant_lng <- read_xlsx("../APECS Master repository/APECS Master repo/ALL_DATA/Lia_fish/Habitat_2019_RAW_2-21-2020.xlsx",
                           sheet = 2)
hab_plant_biom <- read_xlsx("../APECS Master repository/APECS Master repo/ALL_DATA/Lia_fish/Habitat_2019_RAW_2-21-2020.xlsx",
                            sheet = 3)
site <- read.csv("../APECS Master repository/APECS Master repo/ALL_DATA/Site_metadata_allyears.csv", header = TRUE,
                 stringsAsFactors = FALSE)
```

# 2. Kelp data
```{r include=FALSE}
# adjust sites names so that we can convert to universal code
hab_quad$site <- as.factor(hab_quad$site)
levels(hab_quad$site)[levels(hab_quad$site)=="North Fish Egg - kelp"] <- "North Fish Egg-Kelp"
levels(hab_quad$site)[levels(hab_quad$site)=="Baker Island - kelp"] <- "Baker Island - kelp "
unique(hab_quad$site)
# join by quadratID 
kelp_biom <- left_join(hab_plant_biom, hab_quad, by = "quadrat_ID") %>%
  dplyr::select(species_by_quadrat_ID:habitat, date) %>%
  left_join(site.19, by = c("site" = "site_2019")) %>%
  select(c(site, universal_code, species, weight, date))
  
kelp_sum <- kelp_biom %>%
  group_by(universal_code, date, site, species) %>%
  dplyr::summarise(weight_g = sum(weight)) %>%
  mutate(sp = species) %>%
  ungroup() %>%
  dplyr::select(c(sp, universal_code, weight_g, site, date))

kelp_sum$collection_method <- "hand"

kelp <- kelp_sum
kelp <- mutate(kelp, date = ymd(date))
```

# 3 Approximate for 2020
```{r}
kelp_2020 <- kelp %>%
  group_by(sp) %>%
  dplyr::summarise(weight_g_sum = sum(weight_g)) %>%
  mutate(weight_g_by_site = weight_g_sum/5) %>%
  mutate(weight_g_2020 = weight_g_by_site*10) # overestimation of the 6 sites we intend on seining
  
```

# Write out csv
```{r}
write.csv(kelp_2020, "Data/Kelp_sp_estimation_2020.csv")
```

